---
title: "comparison anova"
format: pdf
---

```{r setup, include=FALSE}
#Initial Code Chunk
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(tidyverse)
library(ggthemes)
library(readr)
library(patchwork)
library(car)

diabeted_multiclass <- read_csv("../data/diabeted_multiclass.csv")
```

```{r}
#Data Set Up
data <- diabeted_multiclass %>% 
  select(Diabetes_012,
         BMI)

data <- data %>%
  filter(BMI >= 14.9, BMI <= 33.3)

data$Diabetes_012 <- factor(data$Diabetes_012,
                            levels = c(0, 1, 2),
                            labels = c("no_diabetes", "prediabetes", "diabetes"))
glimpse(data)
```

```{r}
ggplot(data, aes(x = Diabetes_012, y = BMI)) +
  geom_boxplot() +
  labs(title = "BMI Distribution by Diabetes Status",
       x = "",
       y = "BMI") +
  theme_minimal()
```


```{r}
data %>% 
  ggplot()+
  geom_bar(aes(x = Diabetes_012),
           fill="black")+
  theme_few()
```
Differences in sizes of groups. We will make them all the same size to make a balanced test.


```{r}
set.seed(560)
n <- nrow(data[data$Diabetes_012 == "prediabetes", ])

prediabetes <- data[data$Diabetes_012 == "prediabetes", ]
no_diabetes <- slice_sample(data[data$Diabetes_012 == "no_diabetes", ], n=n)
diabetes <- slice_sample(data[data$Diabetes_012 == "diabetes", ], n=n)

test_data <- bind_rows(prediabetes, no_diabetes, diabetes)
```

```{r}
ggplot(test_data, aes(x = Diabetes_012, y = BMI)) +
  geom_boxplot() +
  labs(title = "BMI Distribution by Diabetes Status for Test Data",
       x = "",
       y = "BMI") +
  theme_minimal()
```




### Parametric Anova

```{r}
# The ANOVA MODEl
anova_model <- aov(BMI ~ Diabetes_012, test_data)
summary(anova_model)
```
Anova tells us atleast one groups BMI mean differs from the rest at a very significant P-value

```{r}
# Assumptions Check
# Normal Errors
qqnorm(residuals(anova_model))
qqline(residuals(anova_model), col = "red")
```
Serious Divergence from Normality. 

```{r}
#Assumptions Check
#Homoskadacity
leveneTest(BMI ~ Diabetes_012, data = test_data)
```
Levene's Test reject null that variances are equal. 

One assumptions have be violated. Nonparametric is a great choice.

```{r}
#Post-Hoc Test (as ANOVA is significant)
TukeyHSD(anova_model)
```
all pairwise differences in mean BMI among the three diabetes groups are statistically significant at the 0.05 level. 


Welch's Test does not consider equal variances
```{r}
# Welch's ANOVA
oneway.test(BMI ~ Diabetes_012, data = test_data, var.equal = FALSE)
```
Significant P-value

### Nonparametric - Kruskal Wallis

```{r}
#assumption - same distributions
p_0 <- test_data %>% 
  filter(Diabetes_012 == "no_diabetes") %>% 
  ggplot()+
  geom_bar(aes(x = BMI),
                 fill = "black")+
  theme_few()+
  ggtitle("No Diabetes")

p_1 <- test_data %>% 
  filter(Diabetes_012 == "prediabetes") %>% 
  ggplot()+
  geom_bar(aes(x = BMI),
                 fill = "black")+
  theme_few()+
  ggtitle("Pre-Diabetes")

p_2 <- test_data %>% 
  filter(Diabetes_012 == "diabetes") %>% 
  ggplot()+
  geom_bar(aes(x = BMI),
                 fill = "black")+
  theme_few()+
  ggtitle("Diabetes")

p_0 + p_1 + p_2
```

Same distribution across groups. We are good to proceed.

```{r}
kruskal.test(BMI ~ Diabetes_012, data = test_data)
pairwise.wilcox.test(test_data$BMI, test_data$Diabetes_012, p.adjust.method = "bonferroni")
```

### Smaller Samples

What would happen if we only had 30 observations? Kruskal Wallis is suppose to be good for small samples sizes

```{r}
set.seed(560)
prediabetes_30 <- slice_sample(data[data$Diabetes_012 == "prediabetes", ], n=10)
no_diabetes_30 <- slice_sample(data[data$Diabetes_012 == "no_diabetes", ], n=10)
diabetes_30 <- slice_sample(data[data$Diabetes_012 == "diabetes", ], n=10)

test_data_30 <- bind_rows(prediabetes_30, no_diabetes_30, diabetes_30)
```

```{r}
anova_model_30 <- aov(BMI ~ Diabetes_012, test_data_30)
summary(anova_model_30)
```
fail to reject null


```{r}
qqnorm(residuals(anova_model_30))
qqline(residuals(anova_model_30), col = "red")
```
Errors are normal here.

```{r}
leveneTest(BMI ~ Diabetes_012, data = test_data_30)
```
Fail to Reject Null that variances are equal.

Try Welche's

```{r}
oneway.test(BMI ~ Diabetes_012, data = test_data_30, var.equal = FALSE)
```
Fail to Reject

```{r}
p_0 <- test_data_30 %>% 
  filter(Diabetes_012 == "no_diabetes") %>% 
  ggplot()+
  geom_bar(aes(x = BMI),
                 fill = "black")+
  theme_few()+
  ggtitle("No Diabetes")

p_1 <- test_data_30 %>% 
  filter(Diabetes_012 == "prediabetes") %>% 
  ggplot()+
  geom_bar(aes(x = BMI),
                 fill = "black")+
  theme_few()+
  ggtitle("Pre-Diabetes")

p_2 <- test_data_30 %>% 
  filter(Diabetes_012 == "diabetes") %>% 
  ggplot()+
  geom_bar(aes(x = BMI),
                 fill = "black")+
  theme_few()+
  ggtitle("Diabetes")

p_0 + p_1 + p_2
```

Not really the same distribution

```{r}
kruskal.test(BMI ~ Diabetes_012, data = test_data_30)
pairwise.wilcox.test(test_data_30$BMI, test_data_30$Diabetes_012, p.adjust.method = "bonferroni")
```
Also Fail to Reject here.
